\documentclass{bioinfo}
\copyrightyear{2022}
\pubyear(2022)

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\newcommand\mycommfont[1]{\footnotesize\rmfamily{\it #1}}
\SetCommentSty{mycommfont}
\SetKwComment{Comment}{$\triangleright$\ }{}

\usepackage{natbib}
\bibliographystyle{apalike}

\DeclareMathOperator*{\argmax}{argmax}

\begin{document}
    \firstpage{1}
    \title[Aligning nucleotide sequences with genomap]{genomap: pairwise alignment for nucleotide sequences}
    \author[arfy slowy]{arfy}
    \address{Indonesia}

    \maketitle
    \begin{abstract}
        \section{Motivation:} Recent advances in sequencing technologies promise ultra-long reads of $\sim$100 kilo bases (kb)
        in averange, full-length mRNA or cDNA in heigh troughput and genomic contigs over 100 mega bases (Mb) in length.
        Existing alignment programs are unable or inefficient to process such data at scale, which presses for the development of new
        algorithms.

        \section{Results:} Genomap is a general-purpose alignment program to map DNA or long mRNA sequences againts a large reference database.
        It works with accurate short reads of $\ge$100bp in length, $\ge$1kb genomic reads at error rate $\sim$15\%,
        full-length noisy Direct RNA and cDNA reads, and assembly contigs or closely related full chromosomes of hundreds of megabases in length.
        Genomap does split-read alignment, employs concave gap cost for long insertions and deletions (INDLEs) and introduces new heuristics to reduces
        spurious alignments. It is 3--4 times as fast as mainstream short-read genomic or cDNA mappers at higher accuracy, 
        surpassing most aligners specialized in one type alignment.

        \section{Availability and implementations:}
        \href{https://github.com/slowy07/genomap}

        \section{contact:} slowyarfy3@gmail.com
    \end{abstract}

    \section{Introduction}
    Single Molecule Real-Time (SMRT) sequencing technology and Oxford Nanopore technologies (ONT) produce reads over 10kbp in length at an error rate
    $\sim$15\%. Several aligners have been developed for such
    data~\citep{sedlazeck169557, Chaison:2012aa}. Most of them were five times as slow as mainstream for short-read
    aligners~\citep{Langmead:2012fk} in terms of the number of bases mapped per econd. We speculated there could be subtantial room for speedup on
    the trought that 10kb long sequences should be easier to map than 100bp reads because we can more effectively skip repetitive regions,
    which are often the bottleneck of short-read alignment. We confirmed our speculation by achieving approximate mapping 50 times faster than BWWA-MEM.
    ~\citep{Suzuki:1206aa} extended our work with fast and novel algorithm on generating base-level alignment, which in turn inspired us to develop genomap
    with added functionality.

    Both SMRT and ONT have been applied to the sequencing of spliced mRNAs (RNA-seq). While
    traditional mRNA aligners work~\citep{Wu:2005vn,Iwata:2012aa}, they are not
    optimized for long noisy sequence reads and are tens of times slower than
    dedicated long-read aligners. When developing genomap initially for aligning
    genomic DNA only, we realized minor modifications could enable the base
    algorithm to map mRNAs as well. Genomap becomes a first RNA-seq aligner
    specifically designed for long noisy reads. We have also extended the original
    algorithm to map short reads at a speed faster than several mainstream
    short-read mappers.

    In this article, we will describe the genomap algorithm and its applications
    to different types of input sequences. We will evaluate the performance and
    accuracy of genomap on several simulated and real data sets and demonstrate
    the versatility of genomap.

    \begin{methods}
        \section{Methods}
        Genomap follos a typical seed-chain-align procedure as is used by most full-genome aligners.
        It collects minimalizer~\citep{Roberts:2004fv} of the reference sequences and indexes them in hash table,
        with the key being hash of a minimalizer value being a list of location of the minimalizer copies. Then for each query
        sequence, genomap takess query minimalizers as \emph{seeds}, finds exact matches (i.e \emph{anchors}) to the reference,
        and identifies sets of coliner anchors as \emph{chainss}. If base level aligment is requested, genomap applies dynamic regions
        betwween adjacent anchor in chains.
\end{document}
